### 2016-07-13 paper.tex
Initial commit

% Template for Elsevier CRC journal article
% version 1.2 dated 09 May 2011

% This file (c) 2009-2011 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Engineering

% Changes since version 1.1
% - added "procedia" option compliant with ecrc.sty version 1.2a
%   (makes the layout approximately the same as the Word CRC template)
% - added example for generating copyright line in abstract

%-----------------------------------------------------------------------------------

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          %%
%% Important note on usage                                  %%
%% -----------------------                                  %%
%% This file should normally be compiled with PDFLaTeX      %%
%% Using standard LaTeX should work but may produce clashes %%
%%                                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
%% The 'procedia' option causes ecrc to approximate to the Word template
\documentclass[3p,times,procedia,number]{elsarticle}
\flushbottom

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
%\usepackage{amsmath}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Procedia Engineering}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{K. Kumar, J.-Y. Delenne and K. Soga}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{proeng}

%% Give a short journal name for the dummy logo (if needed)
%\jnltitlelogo{Procedia Engineering}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{cleveref}
\graphicspath{figs/}
% ***************************** Math and SI Units ******************************
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[per=slash]{siunitx} % use this package module for SI units
\usepackage{flexisym}
\usepackage{subcaption}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%\biboptions{authoryear}

 \biboptions{sort&compress}

% if you have landscape tables
\usepackage[figuresright]{rotating}
%\usepackage{harvard}
% put your own definitions here:x
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{1st International Conference on the Material Point Method, MPM 2017}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
%% \dochead can also be used to include a conference title, if directed by the editors
%% e.g. \dochead{17th International Conference on Dynamical Processes in Excited States of Solids}

\title{Transient dynamics of granular piles: continuum v discrete modelling}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[a]{Krishna Kumar\corref{cor1}}
\author[a,b]{Kenichi Soga}
\author[c]{Jean-Yves Delenne}

\address[a]{Department of Engineering, University of Cambridge, Cambridge CB2 1PZ, UK}
\address[b]{Department of Civil Engineering, University of California, Berkeley, USA}
\address[c]{University of Montpellier II, France}

\begin{abstract}
%% Text of abstract
Click here and insert your abstract text.
\end{abstract}

\begin{keyword}
MPM; Granuar dynamics; Tranient flows; Continuum; Discrete Element Modelling;

%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}
\cortext[cor1]{Corresponding author. Tel.: +44-1223 748589.}
\end{frontmatter}

%\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000.}
\email{kks32@cam.ac.uk}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text

%\enlargethispage{-7mm}
\section{Introduction}
\label{sec:intro}
Transient granular flows occur very often in nature. Well-known examples are rockfalls, debris flows,
and aerial and submarine avalanches. They form a major element of lanscape reshape and their
high destructive potential is a substantial factor of risk. Natural granular flows may be triggered 
as a result of different processes such as gradual degradation, induced by weathering or chemical
reactions, liquefaction and external forces such as earthquakes.

Granular flows have been studied in laboratory experiments in different 
geometries such as tilted piles leading to slope failure and surface avalanches 
\citep{Legros2002, Iverson1997} or by considering vertical columns 
of grains collapsing and spreading under their own weight \citep{Lajeunesse2004, 
Lajeunesse2005}. The dynamics observed in the experiments is often nontrivial 
in the sense that the final configurations after the dissipation of the whole kinetic energy  
can not be readily predicted by means of simple laws based on the Mohr-Coulomb nature 
of the material. For example, in collapsing columns, the runout distance 
is found to obey a power-law dependence upon the initial aspect ratio of the 
column.

The observed nontrivial transient dynamics is often correctly 
reproduced by the discrete element method (DEM), which provides 
a powerful tool for the grain-scale analysis of the 
trigger and its subsequent dynamics \citep{Staron2005, Staron2009}. 
However, even in simplified geometries such as those 
investigated in the experiments, the DEM suffers from a serious 
short-coming in the number of particles that can be simulated in 
a reasonnable time. This is a critical issue when more complex geometries or 
long-time granular processes are considered, or when particle size 
distributions are broad. For this reason, most numerical studies 
are performed in 2D or simple particles shapes and size distributions 
are considered.

It is also obvious that classical modeling strategies based on 
the finite element method (FEM) cannot be used for the 
simulation of very large deformations. In various application of FEM, 
this problem is treated by means of technical tools such 
as remeshing. Such mthods are, however, not robust and lead 
to round-off errors and mesh-senstitivity. In contrast, the so-called 
Material Point Method (MPM) is an alternative approach for 
continuum problems that allows for indefinitely large deformations 
without remeshing {Reference??}. In this method, the material points carry the information 
on state variables and a background fixe grid is used to solve the governing equations. 
The information between the material points and the grid is exchanged via 
suitable shape functions. The MPM has been applied with success to 
a number of solid mechanics problems and its theoretical foundations 
have recently been invesitgated by several authors.

In this paper, we are concerned with the ability of the MPM, as a continuum 
approach, to reproduce the evolution of a granular pile 
under its own weight or when destabilized by energy input. In particular, 
a central issue is wheather power-law dependence of the runout distance 
and timewith respect to the initial geometry or energy can be reproduced 
by a simple prescription of the Mohr-Coulomb plastic behavior within a  
MPM code. We therefore perform extensive simulations by varying continuously 
different input parameters and compare the data with those obtained from 
DEM simulations of the same system. We compare in detail the evolution of 
the profile of the pile and its total kinetic energy between the two methods 
and for differerent initial states. As we shall see, the MPM can successfully simulate  
the transient evolution with a single input parameter, namely the 
internal angle of friction. This opens the way to the simulation of 
geological-scale flows on complex topographies.

In the following, we first briefly introduce the MPM  with a Mohr-Coulomb plastic behaviour.  
In section 3, we investigate the collapse dynamics and runout of granular columns. 
We then consider in section 4 the spreading of an initially static pile 
under the action of an impact energy. We conclude with a summary and 
discussion of the most salients results of this work.

\subsection{Numerical set-up}
\label{sec:num}

The DEM sample was composed of $\sim13000$ disks with a uniform distribution of 
diameters by volume fractions ($d_{max} = 1.5 d_{min}$). The mean grain 
diameter and mass are $d\simeq 2.455 $~\si{\mm} and $m\simeq 0.0123$~\si{\kg}, 
respectively. The grains are first poured uniformly into a rectangular box of 
given width and then the right-hand side wall is shifted far to the right 
to allow the grains to spread. A stable granular slope of 13.2~\si{\degree} is 
obtained when all grains come to rest; see~\cref{fig:slope_configuration}. This 
procedure leads to a mean packing fraction $\simeq 0.82$. Soil grains with a 
mean density of 2600~\si{\kg\per\m\cubed} and internal friction coefficient of 
0.4 between grains are considered.

\begin{figure}[tbhp]
  \includegraphics[width=0.95\textwidth]{figs/slope_configuration}
  \caption{Initial geometry and dimensions of the pile subjected to a horizontal 
  excitation.}
  \label{fig:slope_configuration}
\end{figure}

The initial static pile was set into motion by applying a horizontal
gradient velocity $v_{0x}(y) = k (y_{max} - y)$ with $k>0$. The evolution of 
the pile geometry and the total kinetic energy as a function of the initial 
input energy $E_0$ is studied. The run-out distance $L_f$ is the distance of 
the rightmost grain, which is still in contact with the main mass when the pile 
comes to rest. The run-out will be normalised by the initial length $L_0$ of 
the pile, as in the experiments of collapsing columns. The total run-out 
duration $t_f$ is the time taken by the pile to reach its final run-out 
distance $L_f$.

For grain-scale simulations, classical DEM and Contact Dynamics approaches were 
used. This research was done in collaboration with Patrick Mutabaruka, 
University of Montpellier, who performed Contact Dynamics (CD) simulations that 
are presented in this section. A detailed description of the Contact Dynamics 
method can be found in \cite{Moreau1993,Jean1999,Radjai2009,Radjai2011}. 
The CD method is based on implicit time integration of the equations of motion 
and a non-smooth formulation of mutual exclusion and dry friction between 
particles. The CD method requires no elastic repulsive potential and no 
smoothing of the Coulomb friction law for the determination of forces. 
For this reason, the simulations can be performed with large time steps 
compared to discrete element simulations. The unknown variables are particle 
velocities and contact forces, which are calculated at each time step by taking 
into account the conservation of momenta and the constraints due to mutual 
exclusion between particles and the Coulomb friction. An iterative 
algorithm based on a non-linear Gauss-Seidel scheme is used. The only 
contact parameters within the CD method are the friction coefficient $\mu$, the 
normal restitution coefficient $\varepsilon_n$ and the tangential restitution 
coefficient $\varepsilon_t$ between grains.

In MPM simulations, the material point spacing is adopted to be the same as the 
mean grain diameter in DEM. A mesh size of 0.0125m is adopted with 25 material 
points per cell. The effect of mesh size and the number of material points per 
cell is investigated in~\cref{sec:MPM_points_per_cell}. The initial 
configuration of the slope in MPM is shown in~\cref{fig:MPM_Initial_PIC_Slope}. 
Frictional boundary conditions are applied on the left and the bottom 
boundaries by applying constraints to the nodal acceleration. The initial 
vertical stress of the granular pile in equilibrium, before the horizontal 
excitation, in the MPM simulation is shown in~\cref{fig:MPM_Stress_Slope}. The 
distribution of the initial gradient horizontal excitation energy of 50~\si{J} 
on the granular pile is shown in~\cref{fig:MPM_Velocity_Slope}. The 
Mohr-Coulomb model with no dilation is used to simulate the continuum behaviour 
of the granular pile. Periodic shear tests using CD 
(\cref{fig:Sxy_vs_Syy_Slope}), reveals a macroscopic friction coefficient of 
0.22. The evolution of inertial number with friction is presented 
in~\cref{fig:mu_vs_I}.

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[t]{0.75\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figs/MPM_Initial_PIC_Slope}
    \caption{Initial configuration of material points per cell. 25 material points 
    per cell. Cell size of 0.0125m.}
    \label{fig:MPM_Initial_PIC_Slope}
  \end{subfigure} \\
  %
  \begin{subfigure}[t]{0.95\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figs/MPM_Stress_Slope}
    \caption{Initial stress in MPM}
    \label{fig:MPM_Stress_Slope}
  \end{subfigure} \\
  %
  \begin{subfigure}[t]{0.95\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figs/MPM_Velocity_Slope}
    \caption{Initial horizontal velocity for a pile subjected to a horizontal 
    velocity of 50J.}
    \label{fig:MPM_Velocity_Slope}
  \end{subfigure}
  \caption{Initial configuration for the MPM simulation of a pile subjected to 
  horizontal velocities.}
  \label{fig:MPM_Slope_setup}
\end{figure}

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[t]{0.475\textwidth}
    \includegraphics[width=0.95\textwidth]{figs/Sxy_vs_Syy_Slope}
    \caption{Evaluation of the critical state friction angle.}
    \label{fig:Sxy_vs_Syy_Slope}
  \end{subfigure}
  %
  \begin{subfigure}[t]{0.475\textwidth}
    \includegraphics[width=0.95\textwidth]{figs/mu_vs_I}
    \caption{Evolution of Inertial number with friction $\mu$.}
    \label{fig:mu_vs_I}
  \end{subfigure}
  \caption{Periodic shear test using CD~\citep{Mutabaruka2013}.}
  \label{fig:Shear_Test_Slope}
\end{figure}

The natural units of the system 
are the mean grain diameter $d$, the mean grain 
mass $m$ and acceleration due to gravity $g$. For this reason, the length 
scales are normalised by $d$, time by $(d/g)^{1/2}$, velocities by $(gd)^{1/2}$ 
and energies by $mgd$.

%-------------------------------------------------------------------------------
\subsection{Evolution of pile geometry and run-out}
\label{sec:evolution}

\Cref{fig:Gradient_Slope_Profile_200J} shows the initial evolution of the 
granular slope subjected to an initial horizontal energy $E_0 = 61$ (in 
dimensionless units) using MPM. As the granular slope is sheared along the 
bottom, the shear propagates to the top leaving a cavity in the vicinity of the 
left wall. This cavity gets partially filled as the granular mass at the top 
collapse behind the flowing mass due to inertia. Similar behaviour is observed 
during the initial stages of the flow evolution using CD technique 
(\cref{fig:Gradient_Slope_CD_200J}). Due to inertia, the grains at the top 
of the granular heap roll down to fill the cavity, while the pile continues to 
spread.

\begin{figure}[tbph]
  \centering
  \includegraphics[width=\textwidth]{figs/Gradient_Slope_Profile_200J}
  \caption{MPM simulation of the initial stages of granular pile subjected to a 
  gradient horizontal energy.}
  \label{fig:Gradient_Slope_Profile_200J}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/Gradient_Slope_CD_200J}
  \caption{CD simulation of the initial stages of granular pile subjected to a 
  gradient horizontal energy.~\citep{Mutabaruka2013}.}
  \label{fig:Gradient_Slope_CD_200J}
\end{figure}

The flow involves a transient phase with a change in the geometry of the pile 
followed by continuous spreading. The gradient of input energy applied to the 
granular slope mimics a horizontal quake. Despite the creation 
of a cavity behind the flowing mass, the granular heap remains in 
contact with the left wall irrespective of the input 
energy.~\Cref{fig:Runout_Eo_MPM_CD_DEM} shows the normalised run-out distance 
$(L_f - L_0)/L_0$ and total run-out time $t_f$ as a function of the input 
energy $E_0$. Two regimes characterised by a power-law relation between 
the run-out distance and time as a function of $E_0$ can be observed. In the 
first regime, corresponding to the range of low input energies $E_0 < 40 \ 
mgd$, the run-out distance observed varies as $L_f \propto (E_0)^\alpha$ with 
$\alpha 
\simeq 0.206 \pm 0.012$ over nearly one decade. Overall, the run-out distance 
predicted by the continuum approach matches the DEM simulations. At very low 
energies, DEM simulations show longer run-out distance due to local 
fluidisation. The difference in the run-out between DEM and CD arise mainly 
from the scales of description and the inelastic nature of Contact Dynamics. 
Similar behaviour between DEM and CD approaches was observed 
by~\citet{Radjai1997}.

While the run-out exhibits a power-law relation with the initial input energy, 
the DEM simulations show that the flow duration remains constant at a value  
$t_f \simeq 60 \  (d/g)^{0.5}$ irrespective of the value of $E_0$. The constant 
run-out time, in grain-scale simulations, indicates the collapse of grain into 
the cavity left behind the pile. An average run-out speed can be defined as 
$v_s = (L_f - L_0) / t_f$. According to the data, $v_s \propto 
(E_0)^{0.52\pm 0.012}$. The error on the exponent represents the 
error due to the linear fit on the logarithmic scale. Since the initial 
average velocity varies as $v_0 \propto (E_0)^{0.5}$, this difference between 
the values of the exponents suggests that the mobilised mass during run-out 
declines when the input energy is increased.

\begin{figure}[tbph]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Runout_Eo_MPM_CD_DEM}
    \caption{Run-out distance as a function of normalised input kinetic energy.}
    \label{fig:Runout_Eo_MPM_CD_DEM}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Tf_vs_Eo_Slope}
    \caption{Duration of run-out as a function of normalised input kinetic energy.}
    \label{fig:Tf_vs_Eo_Slope}
  \end{subfigure}
  \caption{Evolution of run-out and time as a function of the normalised input 
  energy for a pile subjected a gradient horizontal energy.}
  \label{fig:Slope}
\end{figure}

In the second regime, corresponding to the range of high input energies  $E_0 > 
40 \ mgd$, the run-out distance varies as $L_f \propto (E_0)^{\alpha'}$ over 
one decade with $\alpha' \simeq 0.56\pm 0.04$ while the duration increases as 
$t_f \propto (E_0)^{\beta'}$ with $\beta' \simeq 0.33 \pm 0.02$. Hence, in this 
regime the average run-out speed varies as $v_s \propto (E_0)^{0.498 \pm 
0.01}$. This exponent is close to the value $0.5$ in $v_0 \propto (E_0)^{0.5}$, 
and hence, within the confidence interval of the exponents.
%, in the second 
%regime we may assume $\beta' \simeq \alpha' - 0.5$ and $v_s \propto v_0$. 
In the second regime, both DEM and MPM predict almost the same run-out 
behaviour. However, MPM predicts longer duration with increase in the input 
energy.

It is worth noting that a similar power-law dependence of the run-out distance 
and time are found in the case of granular column collapse with respect to 
the initial aspect ratio. In the column geometry, the grains spread away 
owing to the kinetic energy acquired during gravitational collapse of the 
column.~\citet{Topin2012} found that the run-out distance varies as a power-law 
of the available peak kinetic energy at the end of the free-fall stage with an 
exponent $\simeq 0.5$. This value of exponent is lower than the run-out 
evolution observed in the second regime. This is, however, physically plausible 
since the distribution of kinetic energies at the end of the collapse 
is more chaotic than in this case where the energy is supplied from the very 
beginning in a well-defined shear mode. As pointed out by~\citet{Staron2005}, 
the distribution of kinetic energies is an essential factor for the run-out 
distance.

%-------------------------------------------------------------------------------------------
\subsection{Decay of kinetic energy}
\label{sec:decay}

The non-trivial evolution of the pile geometry in two regimes suggests that 
the energy supplied to the pile is not simply dissipated by shear and friction 
along the bottom plane. It is important to split the kinetic energy into 
vertical and horizontal components ($K_{Ex}$ and $K_{Ey}$) of the velocity 
field. Although, the input energy is in the $x$ component, a fraction of 
the energy is transferred to the vertical component of the velocity field and 
dissipated during the transient phase. The evolution of kinetic energy is 
studied to understand the behaviour of granular flow that is consistent with 
the evolution of the pile shape.

The evolution of total kinetic energies $E_k$ with time for different values of 
the input energy $E_{ki}$ based on MPM simulations are shown 
in~\cref{fig:Energy_Time_Slope}. The MPM simulation shows two distinct regimes 
in the normalised kinetic energy plot as a function of normalised time 
in~\cref{fig:Normalised_Energy_Time_Slope}. However, the DEM simulations 
(\cref{fig:Normalised_Energy_Time_Slope_DEM}) show that the energy evolution 
corresponding to the low energy regime nearly collapse on to a single time 
evolution. This is consistent with the observation of run-out time $t_f$ being 
independent of the input energy. In contrast, MPM simulations predict a power 
law relation between the run-out duration and input energy. However, the 
plots corresponding to the high energy regime (\cref{fig:Energy_Time_Slope}), 
collapse only at the beginning of the run-out i.e. for $t < t_1 \simeq 7.5 \ 
(d/g)^{0.5}$. Although MPM simulations show a longer duration of run-out 
(\cref{fig:Energy_Time_Slope}), the total kinetic energy is completely 
dissipated at $t = 60 \sqrt{d/g}$. DEM simulations predict $t = 80 \sqrt{d/g}$ 
for the kinetic energy to be completely dissipated. This is due to grain 
rearrangement at the free surface (\cref{fig:voro_500}). The granular mass 
densifies as the flow progresses, after the initial dilation phase for $t = 
20\sqrt{d/g}$.

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \includegraphics[width=\textwidth]{figs/Energy_Slope}
    \caption{Evolution of total kinetic energy with time.}
    \label{fig:energy_slope}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Normalised_Energy_Time_Slope}
    \caption{Evolution of normalised kinetic energy with normalised time.}
    \label{fig:Normalised_Energy_Time_Slope}
  \end{subfigure}
  \caption{Evolution of kinetic energy with time (MPM) for a pile subjected to 
  gradient input velocities.}
  \label{fig:Energy_Time_Slope}
\end{figure}

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=0.9\textwidth]{figs/Normalised_Energy_Time_Slope_DEM}
  \caption{Evolution of normalised kinetic energy with normalised time for a pile 
  subjected to gradient input velocities.}
  \label{fig:Normalised_Energy_Time_Slope_DEM}
\end{figure}

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=0.9\textwidth]{figs/voro_500}
  \caption{Evolution of packing density with time $E_0 = 152 mgd$ (DEM).}
  \label{fig:voro_500}
\end{figure}

\Cref{fig:Normalised_KEx_KEy_Slope} displays the evolution of kinetic energy 
in the translational ($E_x$ and $E_y$) degrees of freedom. $E_x$ decays similar 
to the total energy dissipation, but $E_y$ increases and passes through a peak 
before decaying rapidly to a negligible level. The transient is best 
observed for $E_y$, which has significant values only for $t< t_1$. This energy 
represents the proportion of kinetic energy transferred to the $y$ component of 
the velocity field  due to the destabilisation of the pile and collapse of 
grains in the cavity behind the pile. Higher proportion of vertical 
acceleration $E_{ky}/E_0$ is observed for lower values of input energy $E_0$. 
This means that, at lower input energies a larger fraction of the energy is 
consumed in the destabilisation process. Whereas at a higher input energies, 
most of the energy is dissipated in the spreading phase. For this reason, the 
total duration $t_1$ of this destabilisation phase is nearly the same in 
both regimes and its value is controlled by gravity rather than the input 
energy. The height of the pile being of the order of $80 \ d$, the total 
free-fall time for a particle located at this height is $\simeq 12 \ 
(d/g)^{0.5}$, which is of the same order as $t_1$. DEM simulations show that 
the contribution of the rotational energy during the transient stage and the 
spreading stage is negligible.

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \includegraphics[width=\textwidth]{figs/Normalised_KEx_Slope}
    \caption{Evolution of normalised horizontal kinetic energy with time.}
    \label{fig:Normalised_KEx_Slope}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Normalised_KEy_Slope}
    \caption{Evolution of normalised vertical kinetic energy with time.}
    \label{fig:Normalised_KEy_Slope}
  \end{subfigure}
  \caption{Evolution of vertical and horizontal kinetic energy with time (MPM) 
  for a pile subjected to gradient input velocities.}
  \label{fig:Normalised_KEx_KEy_Slope}
\end{figure}

To analyse the second phase for higher input energies, the 
kinetic energy $E'_{kx0}$ at the end of the transient phase is 
considered. This energy is responsible for most of the run-out, hence it is 
expected to 
control the run-out distance and time.~\Cref{fig:Normalised_KExExop_Slope} 
shows the evolution of $E_{kx}$ normalised by $E'_{kx0}$ as a function of time. 
The plots have seemingly the same aspect but they show different decay times. A 
decay time $\tau$ can be defined as the time required for $E_{kx}$ to decline 
by a factor $1/2$.~\Cref{fig:ExEx0_vs_ttau} shows the same data in 
which the time $t'$ elapsed since $t_1$, normalised by $\tau$. Interestingly, 
now all the data nicely collapse on to a single curve. However, this curve can 
not be fitted by simple functional forms such as variants of exponential decay. 
This means that the spreading of the pile is not a self-similar process in 
agreement with the fact that the energy fades away in a finite time $t'_f$.

\clearpage

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\textwidth]{figs/Normalised_KExExop_Slope}
  \caption[Evolution of the normalised horizontal kinetic energy as function of 
  time since the transient phase.]{Evolution of kinetic energy in the $x$ 
  component of the velocity field normalised by the available kinetic energy at 
  the
  end of the transient phase as a function of time elapsed since the
  same instant (MPM).}
  \label{fig:Normalised_KExExop_Slope}
\end{figure}

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\textwidth]{figs/EkxKoTTau_Slope}
  \caption[Evolution of the normalised horizontal kinetic energy as function of 
  the normalised time since the transient phase.]{Evolution of kinetic energy in 
  the $x$ component of 
  the velocity field  normalised by the available kinetic energy at the end of 
  the transient as a function of normalised time (MPM).}
  \label{fig:ExEx0_vs_ttau}
\end{figure}

The scaling of the data with the decay time $\tau$ suggests that the 
run-out time, since the beginning of the second phase, $t'_f$ might be a simple 
function of $\tau$.~\Cref{fig:tp_tau_mgd} shows both $t'_f$ and $\tau$ as 
a function of $E'_{x0}$, where a power-law relation can be observed for both 
time scales. The run-out time $t'_f \propto (E'_{x0})^{\beta'}$ has the 
same exponent $\beta' \simeq 0.33 \pm 0.02$ as $t_f$ as a function of $E_0$. 
For the decay time we have $\tau \propto (E'_{x0})^{\beta''}$ with $\beta'' 
\simeq 0.38 \pm 0.03$. The relation between the two times can thus be expressed 
as (\cref{fig:tp_tau})
\begin{equation}
  t'_f = k  \ \tau \, (E'_{x0})^{\beta'' - \beta'} \,,
  \label{eqn:t'f}
\end{equation}
where $k \simeq 5.0 \pm 0.4$ and $\beta'' - \beta' \simeq 0.05 \pm 0.05$. This 
value is small enough to be neglected. It is therefore plausible to assume that 
the run-out time is a multiple of the decay time and the spreading process is 
controlled by a single time. A weak dependence on the energy $E'_{kx0}$ is 
consistent with the fact that the energy available at the beginning of the 
second phase is not dissipated in the spreading process (calculated from the 
position of the tip of the pile) since the pile keeps deforming by the 
movements of the grains at the free surface even when the tip comes to rest. 
This can explain the small difference between the two exponents as observed 
here.

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.92\textwidth}
    \includegraphics[width=\textwidth]{figs/tp_tau_mgd}
    \caption{Power law evolution of $t'_f$ and $\tau$ as a function of kinetic 
    energy $E'_{kx0}$.}
    \label{fig:tp_tau_mgd}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.92\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/tp_tau}
    \caption{Linear relationship between decay time and run-out time after the 
    transient as a function of the normalised kinetic energy $E_{kx0}$.}
    \label{fig:tp_tau}
  \end{subfigure}
  \caption{Decay time and run-out time as a function of the normalised kinetic 
  energy $E_{kx0}$.}
  \label{fig:tp_Tau}
\end{figure}

%-------------------------------------------------------------------------------
\subsection{Effect of friction}
\label{sec:parameters}

The run-out distance, duration of flow, and the dissipation of kinetic energy 
are controlled by the input energy and collective dynamics of the whole pile. 
However, the run-out behaviour is also expected to depend on the base friction. 
A series of simulations with different values of base friction was performed 
using MPM to analyse the influence of friction on the run-out behaviour. The 
influence of friction on the run-out behaviour for different input energies is 
shown in~\cref{fig:runout_fric_slope}. The run-out distance decreases with 
increase in the basal friction. The exponent of the 
power-law relation between the run-out and input energy has a weak dependence 
on the base friction, however, the proportionality constant is affected by the 
change in the base friction. This behaviour is similar to that observed in 
granular column collapse with varying initial 
properties~\citep{Balmforth2005,Lajeunesse2005}.

CD simulations using different values of coefficient of restitution show no 
difference in the run-out behaviour. At large input energies, the pile remains 
in a dense state so that multiple collisions inside the pile occur at small 
time scales compared to the deformation time. When the restitution coefficients 
are increased, more collisions occur during a longer time interval but the 
overall energy dissipation rate by collisions remains the same. This effect is 
a seminal example of collective effects which erase the influence of local 
parameters at the macroscopic scale.

In contrast to the restitution coefficients, the effect of friction 
coefficient is quite important for the run-out. MPM simulations with 
varying friction coefficient show that, both the run-out distance and the 
decay time decrease as the friction coefficient is increased. This 
effect is much more pronounced at low values of the friction coefficient. 
The run-out time, for example, is reduced by a factor of approximately 4 as 
$\mu_s$ is increased from 0.1 to 0.2 while the change in the run-out and 
duration is less affected with increase in friction coefficient. This 
``saturation effect'' can be observed in a systematic way in simple shear 
tests. The dissipation rate may reach a saturation point where the dilation of 
granular materials and rolling of the grains change in response to increase in 
friction coefficient~\citep{Estrada2008}.

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/runout_fric_slope}
    \caption{Effect of friction on the run-out distance}
    \label{fig:runout_fric_slope}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/time_fric_slope}
    \caption{Effect of friction on the duration of run-out.}
    \label{fig:time_fric_slope}
  \end{subfigure}
  \caption{MPM simulations of effect of friction on the run-out behaviour of 
  slopes subjected to horizontal excitation.}
  \label{fig:fric_slope}
\end{figure}

\section{Conclusion}
Natural granular flows are triggered by different mechanisms. The distribution 
of kinetic energy in the granular mass is found to have an effect on the flow 
kinematics. A multi-scale analyses of a granular slope subjected to horizontal 
velocities are performed and the following conclusions are derived:

\begin{itemize}

  \item A power-law dependence of the run-out distance and time as a 
    function of the input energy is observed. The power-law behaviour is found to 
    be a generic feature of granular flow dynamics.

  \item The values of the power-law exponents are not simple functions of the 
    geometry.

  \item Two regimes with different values of the exponents: 
    a low-energy regime and a high-energy regime are observed.

  \item The low energy regime reflects mainly the destabilisation of the pile, 
    with a run-out time independent of the input energy.

  \item The second regime is governed by the spreading dynamics 
    induced by higher input energy. The evolution of granular slope in the 
    high-energy regime can be described by a characteristic decay time, which is 
    the time required for the input energy to decay by a factor of 0.5.

  \item The run-out distance and the decay time decrease as the friction 
    increases. This effect is much more pronounced at low values of friction.

  \item  At low input energy, the distribution of kinetic energy in the system is 
    found have a significant effect on the run-out, as the energy is mostly 
    consumed in the destabilisation process.

  \item At higher input energy, where most of the energy is dissipated during 
    the spreading phase, the run-out distance has a weak dependence on the 
    distribution of velocity in the granular mass.

  \item The duration of the flow shows similar behaviour to the run-out, however, 
    a slope subjected to a gradient velocity flows quicker than a slope subjected 
    to a uniform horizontal velocity.

  \item The material property and the distribution of kinetic energy in the 
    system has a non-trivial influence on the flow kinematics and the internal flow 
    structure.

  \item MPM is successfully able to simulate the transient evolution with a 
    single input parameter, the macroscopic friction angle.

\end{itemize}

This study exemplifies the ability of MPM, a continuum approach,  in modelling 
complex granular flow dynamics and opens the possibility of realistic 
simulation of geological-scale flows on complex topographies.

\section*{Acknowledgements}

\bibliographystyle{model1-num-names}
\bibliography{references} 
\end{document}

%%
%% End of file `procs-template.tex'.